---
title: "Extracting Ouderbijdrage"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

# Extracting the Ouderbijdrage

In many schools, they request a voluntary ouderbijdrage that varies significantly from school
to school. This information is only present in the text of the schoolgids.

Our goal is to find the value of the ouderbijdrage in each of the school gids.

An initial review of schoolgids suggests that identifying this information is best
done at the sentence level. So we will try to develop a probablistic model that can identify
the sentences which specify the ouderbijdrag.


## Libraries

```{r}
library(e1071)
library(tm)
```

## Downloading the data

We will download a sub-sample of school guides that have tokenized by frog, an advanced natural language
toolkit for Dutch.

```{r}
datafile <- "schoolgids2017v4_frogged_200.rds"
if(!file.exists(datafile)) {
  download.file("https://storage.googleapis.com/schoolgids/schoolgids2017v4/schoolgids2017v4_frogged_200.rds", datafile)
}

tokens <- readRDS(datafile)
tokens$sent_id <- paste(tokens$school, tokens$sent, sep="")
```

## Find all the sentences with numbers

After reviewing a random sample of schoolgids PDFs, we are confident that the ouderbijdrage is always expressed as a numeral, 
for example "40 euro" and not "vijfig euro", so we will limit our analysis to sentences which contain a token that is a 
telwoord ('TW') and a numeral:

```{r}

# Compute the numeric value of each telwoord token, handling "," as decimal separators.
tokens$number <- ifelse(tokens$pos == 'TW', 
                         as.numeric(gsub(tokens$word, pattern = ",", replacement = ".")),
                         NA)

sentences_with_numbers <- unique(tokens$sent_id[!is.na(tokens$number)])

tokens <- subset(tokens, sent_id %in% sentences_with_numbers)

```


## Building a training sample

First we need to label our dataset. We'll manually review the PDFs to find the sentences that contain
the ouderbijdrage as a numeric value. We sampled 20 schools from the 200 and divided the PDFs up for manual
review in [Google Spreadsheet](https://docs.google.com/spreadsheets/d/1E5VmiCcuTqbkk1C3ropHuZAvPEJnUn2kNsr6i5FRXV0/edit#gid=0)

Once the "correct" sentences were identifed that contain the desired information, we load the table back in.

```{r}
labels <- read.delim("labels.tab", col.names = c("school", "correct_sent"), stringsAsFactors = FALSE)
```

This table a row for each of the schools we reviewed, and the index of the "correct" sentance that
contains the ouder bijdrag.

```{r,echo=FALSE}
knitr::kable(labels)
```

Before building our model, we will review the unigrams and bigrams that are found in the sentences that
we labeled as containing the ouderbijdrage. This should provide suggestions as to which features to include
in our model.

```{r}

# merge the tokens with our labels
labeled_tokens <- merge(tokens, labels)
labeled_tokens$correct <- labeled_tokens$sent == labeled_tokens$correct_sent

# find the token rows that were in the flagged sentences
correct_sentences <- subset(labeled_tokens, sent == correct_sent )

# find the most frequent unigrams
unigrams <- correct_sentences$lemma
head(sort(table(unigrams), decreasing = TRUE), n = 30)


# find the most frequent bigrams
bigrams <- unlist(tapply(correct_sentences$lemma, correct_sentences$sent_id, function(tokens) {
  vapply(ngrams(tokens, 2L), paste, "", collapse = " ")
}))
head(sort(table(bigrams), decreasing = TRUE), n = 10)

```

## Build the feature matrix

Using the ideas from our labeled dataset, we will build a feature matrix that includes all
sentences with a number in the rows, and features of those sentences in the columns.

```{r}

# Split the table of tokens into one table per sentence.
# This makes it easier to compute features for each sentance using sapply()

sent_list <- split(tokens, tokens$sent_id)

feature_matrix <- data.frame(
  sent_id = names(sent_list),
  school = sapply(sent_list, function(s) s$school[1]),
  sent = sapply(sent_list, function(s) s$sent[1]),
  per_kind = sapply(sent_list, function(s) grepl(pattern="per kind", paste(s$lemma, collapse=" "))),
  school_year = sapply(sent_list, function(s) grepl(pattern="2017.2018", paste(s$lemma, collapse=" ")))
)

for(unigram in c("vaststellen", "verenigingsbijdrage", "€", "ouderbijdrage", "contributie", "bedrag")) {
  feature_matrix[[unigram]] <- sapply(sent_list, function(s) unigram %in% s$lemma)
}

# Keep track of our features, which exclude the first three
# columns containing school and sentance metadata

features <- names(feature_matrix)[-(1:3)]
```

```{r, echo=FALSE}
knitr::kable(feature_matrix[1:10, ], row.names = FALSE)
```

Now we're ready to build our training matrix, that includes only the sentences (rows) from the schools that we 
manually labled.

```{r}

labeled_matrix <- merge(feature_matrix, labels, all.x = TRUE)

training_matrix <- subset(labeled_matrix, !is.na(correct_sent))
training_matrix$correct <- training_matrix$sent == training_matrix$correct_sent

summary(training_matrix)
```

We now use the svm algorithm to build model which can predict which sentences contain
the ouderbijdrage.

```{r}

model <- svm(training_matrix[, features], as.factor(training_matrix$correct))
```

Finally, we can use this to predict which of the sentences in the rest of the corpus
contain the ouderbijdrage, and print the list of sentences that have been identified.

```{r}
matching <- predict(model, feature_matrix[, features]) == "TRUE"

result <- data.frame(
  school = feature_matrix$school[matching],
  text = sapply(sent_list[matching], function(s) paste(s$word, collapse = " "))
)

knitr::kable(result, row.names = FALSE)

```

