---
title: "SVM and cluster analysis in school terms"
output: github_document
---

* We will continue with the school as the unit of analysis and build on the term-frequency matrices we looked at last week and introduce two machine learning techniques:
  
  - **Supervised machine learning** works by "training" a model on a set of "labeled" documents.
  
  - **Clustering** is an unsupervised technique to uncover groups in the datasets

## Supervised machine learning

* Most machine learning algorithms require an input set of features

* We need to transform the school guides into a matrix of schools and factors

* Term-frequency matrices are one example of features

* Machine learning algorithms can be very accurate classifiers, they tend to have low explanatory power.

## Training 

Load the full(-ish) the pre-computed normalized terms matrix from last week:

```{r}
ndtm <- readRDS("schoolgids2017v2_ndtm_v2.rds")
schools <- read.csv("../schools.csv", sep=";", stringsAsFactors = FALSE)
ndtm_meta <- merge(data.frame(VESTIGINGSNUMMER = rownames(ndtm)), schools, all.x = TRUE)
```

The `ndtm` matrix contains a normalized term-frequent matrix for all the schools in our corpus.

```{r}
ndtm[1:10, 1:20]
```

* We will split our sample into a training and testing samples

```{r}
training <- sample(nrow(ndtm), size = 500)

predictors <- ndtm[training, ] 
response <- as.factor(ndtm_meta$DENOMINATIE[training])
  
library(e1071)
model <- svm(x = predictors, y = response) 
```

* Now we can use this model to predict the schools in the test set 


```{r}

predicted <- predict(model, ndtm[-training, ])

table(predicted)
```

* Let's see how our model did:

```{r}

correct <- predicted == ndtm_meta$DENOMINATIE[-training]
table(correct)
```

Let's look at how we did by category:

```{r}

table(ndtm_meta$DENOMINATIE[-training], correct)
```